Developing a preprint culture in biology

### 2012-10-02 ms.tex

Public preprint servers allow authors to make manuscripts publicly available
before, or in parallel to, submitting them to journals for traditional
peer-review. The rationale for preprint servers is fundamentally simple: to make
the results available to the scientific community as soon as possible rather
than to wait until the peer-review process is fully completed. The goal of arXiv
and open preprint servers is not to avoid the peer-review process; almost all
manuscripts on arXiv are submitted to peer-review.  Sharing ideas as quickly as
possible using preprint servers has numerous advantages.  These include rapid
dissemination of work in progress to a wider audience, including readers in
developing countries where access to subscription journals is often a limiting
factor in scientific research.  Furthermore, preprints provide the opportunity
to solicit feedback from a larger pool of reviewers and can be seen as an
integral aspect of a vigorous peer-review process \cite{hoc12}.  Preprint
servers therefore increase the number of opportunities for review and revision
prior to publication, resulting in higher quality submissions and also could
alleviate reviewer burden.

Preprints gained popularity 20 years ago with the advent of arXiv, an open
preprint server widely used in physics and mathematics \cite{gin11}. Preprints
are also integral to the culture of other scientific fields.  Paul Krugman noted
that, in economics, the \emph{traditional model of submit, get refereed,
publish, and then people will read your work broke down a long time ago. In
fact, it had more or less fallen apart by the early 80s} \cite{kru12}. In
addition to a section in arXiv, economists have also the RePEc (Research Papers
in Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.  Despite the success of this approach in other
fields, most manuscripts in biology are not submitted to a public archive before
being submitted to peer-review.  In this article, we will first highlight the
advantages of open preprint servers for both scientists and publishers. We will
also debunk a few misconceptions, discuss the policies of major publishers in
biology, and briefly review the most popular open preprint servers.

@@

Submitting to a preprint server is becoming more common in biology, even though
it still involves a minority of papers. The quantitative biology section in
arXiv is experiencing faster growth in submissions than any other field
\cite{cal12}. Most scientific journals are preprint-friendly: Nature, PLOS, BMC,
PNAS, Science (mostly) \ref{table:policies}, and all the journals from Elsevier
and Springer.  The Ecological Society of America and the Genetics Society of
America recently changed their policies to allow public preprints.  Few
scientific publications will not consider a manuscript submitted to a public
preprint server.  Still, a few journals adopt a ``by default'' hostile attitude
towards preprints, mostly due to the lack of clear policy of the publishers, or
perhaps because a preprint culture has not developed in biology and the practice
is still considered unusual. As an example, Wiley-Blackwell, which publishes
some leading journals, has no official policy on the subject
\ref{table:policies}.

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new publishing system
that provides both a preprint server, and a peer reviewed journal.  It is
focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a home in biology.  PeerJ allows
commenting on posted preprints, improving the potential for pre-publication
dialog. In addition, preprints can be made private if the authors choose, and
shared only with selected colleagues. While this reduces some of the benefits of
preprints described above, it may allow some researchers who would not otherwise
post preprints to begin to explore the possibility in a manner appropriate to
their current circumstances.

@@

\subsection{Github et al.?}

This manuscript was developed entirely as an open project on github. github is
one of several hosting services for collaborative development using the git
revision control software.  git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. It allows the users to share code, track changes, and merge
versions. For example, during the development of this manuscript, one author
would fork the project (\emph{i.e.} make a personal copy), modify it, and then the
changes were merged back into the main branch. These techniques are widely used
in open source software development. It goes even farther than submitting
preprints by opening the entire writing process.

@@

Preprint servers also facilitate communication between disciplines, bridging
cultural as well as geographic divides. Examples include a recent series of
papers on the theory of natural selection that was posted to arXiv
simultaneously with its publication in the \emph{Journal of Evolutionary
Biology} \cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}. Other
submissions in this category include evolutionary and ecological theory by
authors trained in physics and computer science.  Since authors in these fields
regularly check arXiv, submitting preprints may be the most effective way
biologists can help others avoid ``repeated work'' \cite{de2011contribution} and
form a synthetic community of evolutionary theorists from disparate backgrounds.
The advantages are clear and the costs are low.

>>>>

Public preprint servers allow authors to make manuscripts publicly available
before, or in parallel to, submitting them to journals for traditional
peer-review. The rationale for preprint servers is fundamentally simple: to make
the results available to the scientific community as soon as possible rather
than to wait until the peer-review process is fully completed. The goal of arXiv
and open preprint servers is not to circumvent the peer-review process; almost all
manuscripts on arXiv are submitted to peer-review.  Sharing ideas as quickly as
possible using preprint servers has numerous advantages.  These include rapid
dissemination of work-in-progress to a wider audience, including readers in
developing countries where access to subscription journals is often a limiting
factor in scientific research.  Furthermore, preprints provide the opportunity
to solicit feedback from a larger pool of reviewers and can be seen as an
integral aspect of a vigorous peer-review process \cite{hoc12}.  Preprint
servers therefore increase the number of opportunities for review and revision
prior to publication, resulting in higher quality submissions and also could
alleviate reviewer burden.

Preprints gained popularity 20 years ago with the advent of arXiv, an open
preprint server widely used in physics and mathematics \cite{gin11}. Preprints
are also integral to the culture of other scientific fields.  Paul Krugman noted
that, in economics, the \emph{traditional model of submit, get refereed,
publish, and then people will read your work broke down a long time ago. In
fact, it had more or less fallen apart by the early 80s} \cite{kru12}. In
addition to a section in arXiv, economists have also the RePEc (Research Papers
in Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.  Despite the success of this approach in other
fields, most manuscripts in biology are not submitted to a public archive before
being submitted to peer-review.  In this article, we first highlight the
advantages of open preprint servers for both scientists and publishers. We
then debunk a few misconceptions, discuss the policies of major publishers in
biology, and briefly review the most popular open preprint servers currently available.

@@

Submitting to a preprint server is becoming more common in biology, even though
it still involves a minority of papers. The quantitative biology section in
arXiv is experiencing faster growth in submissions than any other field
\cite{cal12}. Most scientific journals are preprint-friendly: Nature, PLOS, BMC,
PNAS, Science (mostly) \ref{table:policies}, and all the journals from Elsevier %something is off in the formatting here
and Springer.  The Ecological Society of America and the Genetics Society of
America recently changed their policies to allow public preprints.  Few
scientific publications will not consider a manuscript submitted to a public
preprint server.  Still, a few journals adopt a ``by default'' hostile attitude
towards preprints, mostly due to the lack of clear policy of the publishers, or
perhaps because a preprint culture has not developed in biology and the practice
is still considered unusual. As an example, Wiley-Blackwell, which publishes
some leading journals, has no official policy on the subject
\ref{table:policies}.

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new publishing system
that combines both a preprint server, and a peer reviewed journal.  It is
focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a home in biology.  PeerJ allows
commenting on posted preprints, improving the potential for pre-publication
dialog. In addition, preprints can be made private if the authors choose, and
shared only with selected colleagues. While this reduces some of the benefits of
preprints described above, it may allow some researchers who would not otherwise
post preprints to begin to explore the possibility in a manner appropriate to
their current circumstances.

@@

\subsection{Github}

This manuscript was developed entirely as an open project on github. github is
one of several hosting services for collaborative development using the git
version control system (VCS).  git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous contributers to work asynchronously on the same project, often in parallel branches, all of which can be effortlessly merged and version controlled. Although created primarily for software development, git is ideal for academic research since it provides a way to version control and collaborate on every step of the manuscript development process, from data manipulation and analysis to writing and revision. For example, during the development of this manuscript, each author
would clone the project (\emph{i.e.} make a personal copy), modify it, and then merge their changes into a master branch.This takes the preprint process to an entirely new level, where the entire writing process is open from the beginning.
% I would avoid using the term forking since it is not a central piece of the section. It is one specific way of collaborating on a github repo when everyone does not have rights to it. People could just as easily work on a single branch, all with write access, either directly to the master branch or to individual branches that could be merged. So in other words, we should keep the language general so people don't get bogged down with trying to understand what a fork is.
% also I changed one author clones, then submits pull request langauge a bit since the main point of using git is that we can all work simultaneously unlike svn.
% I took out "These techniques are widely used in open source software development." because git can also be used for closed source software, offering the same advantages, where the repositories are private. Paid git accounts allow for private repos.

@@

% Aren't we bridging more than cultural and geographic? I mean, we are essentially cross pollinating ideas. e.g. It took a lot time for MCMC to become popular in ecology even though it was widely used in other disciplines a decade or two earlier.
Preprint servers also facilitate communication between disciplines, bridging
cultural as well as geographic divides. Examples include a recent series of
papers on the theory of natural selection that was posted to arXiv
simultaneously with its publication in the \emph{Journal of Evolutionary
Biology} \cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}. Other
submissions in this category include evolutionary and ecological theory by
authors trained in physics and computer science.  Since authors in these fields
regularly check arXiv, submitting preprints may be the most effective way
biologists can help others avoid ``repeated work'' \cite{de2011contribution} and
form a synthetic community of evolutionary theorists from disparate backgrounds.
The advantages are clear and the costs are low.
